### 2021-04-21 population_genetics/population_genetics.Rmd

---
layout: default
title: "Population Genetics"
nav_order: 4
has_children: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Population Genetics

Check out [Quality Control and Filtering](https://marineomics.github.io/Template_Site_1/Quality_Control_and_Filtering.html)

General introduction to whole genome sequencing and reduced representation sequencing. Pros and Cons for each.

``````{r, echo=FALSE,eval=FALSE}
library(knitr)
my.jekyll.site <- "marineomics.github.io/Template_Site_1/"
KnitPost <- function(input, base.url = my.jekyll.site) {
  opts_knit$set(base.url = base.url)
  fig.path <- paste0("images/", sub(".Rmd$", "", basename(input)), "/")
  opts_chunk$set(fig.path = fig.path)
  opts_chunk$set(fig.cap = "center")
  render_jekyll()
  knit(input, envir = parent.frame())
}

setwd("C:/Users/jmcgirr/Documents/GitHub/MarineOmics/Template_Site_1.github.io/population_genetics/")
KnitPost("C:/Users/jmcgirr/Documents/GitHub/MarineOmics/Template_Site_1.github.io/population_genetics/population_genetics.Rmd")

```