### 2021-05-03 about/about.Rmd
Updated Rmd files

---
layout: default
title: "About"
nav_order: 2
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# About

Welcome! The MarineOmics working group is part of the RCN for Evolution in Changing Seas, and was formed to promote reproducible and robust genomic research in evolutionary biology and marine science. Methods in next generation sequencing and bioinformatics advance at a pace that quickly render any static set of best practices as obsolete. To support robust ‘omics research now and moving forward, the working group will disseminate guidelines through an easily-interfaced and dynamic website that will be deployed in August 2021. To help develop content for the website, we are hosting Q & A panel seminars with experts on various topics relating to ‘omics analyses in non-model systems.

The working group is also currently hosting three undergraduate research interns with funding from the RCN-ECS. These interns are part of a subgroup conducting a literature survey to characterize trends in data availability, code accessibility, and methods reproducibility for a peer-reviewed article.

``````{r, echo=FALSE,eval=FALSE}
library(knitr)
my.jekyll.site <- "marineomics.github.io"
KnitPost <- function(input, base.url = my.jekyll.site) {
  opts_knit$set(base.url = base.url)
  fig.path <- paste0("images/", sub(".Rmd$", "", basename(input)), "/")
  opts_chunk$set(fig.path = fig.path)
  opts_chunk$set(fig.cap = "center")
  render_jekyll()
  knit(input, envir = parent.frame())
}

setwd("~/Postdoc_Projects/RCN_MarineOmics/Website/marineomics.github.io/about/")
KnitPost("~/Postdoc_Projects/RCN_MarineOmics/Website/marineomics.github.io/about/about.Rmd")

```