### 2021-10-26 guidelines/guideline-intro.Rmd

---
layout: default
title: Best Principles
editor_options: 
  chunk_output_type: console
output:
  html_document
has_children: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# This content is still under construction!

# Best Principles for 'Omics in Nonmodel Systems

What are 'omics methods?

Best principles vs Best Practices.

Here are a collection of guidelines, tutorials, recommended readings, and method evaluations for various 'omics analyses that are commonly used in studies of nonmodel systems (particularly marine species).

## Omics Approaches {.tabset .tabset-fade .tabset-pills}
### Population Genomics

* [Choosing a population genomics sequencing approach](WGS_intro.md)  
* Whole Genome Resequencing  
* Reduced Representation Sequencing

### Functional Genomics

* Transcriptomics with RNA-Seq  
* Differential gene expression

## {-}

```{r, echo=FALSE,eval=FALSE}
library(knitr)
my.jekyll.site <- "marineomics.github.io"
KnitPost <- function(input, base.url = my.jekyll.site) {
  opts_knit$set(base.url = base.url)
  fig.path <- paste0("images/", sub(".Rmd$", "", basename(input)), "/")
  opts_chunk$set(fig.path = fig.path)
  opts_chunk$set(fig.cap = "center")
  render_jekyll()
  knit(input, envir = parent.frame())
}

setwd("~/Postdoc_Projects/RCN_MarineOmics/Website/marineomics.github.io/guidelines")
KnitPost("~/Postdoc_Projects/RCN_MarineOmics/Website/marineomics.github.io/guidelines/guideline-intro.Rmd")

```