### 2021-11-04 TOC.Rmd

---
title: Table of contents
---

- **[Panel Seminars](panels.html)**

- **Guidelines for Omics Analyses**
    - [Choosing a Population Genomics Approach](popgen_intro.html)
    - ](Population_Strata.html)
    - [Locus-based statistics and missing data](Locus_Stats.html)
    - [Genotypic evenness, richness, and diversity](Genotypic_EvenRichDiv.html)
    - [Linkage disequilibrium](Linkage_disequilibrium.html)
    - [Population structure](Pop_Structure.html)
    - [Minimum Spanning Networks](Minimum_Spanning_Networks.html)
    - [AMOVA](AMOVA.html)
    - [Discriminant analysis of principal components (DAPC)](DAPC.html)

- **Part III: Using genomic data in population genetics**
    - [Population genomics based on high throughut sequencing (HTS)](intro_vcf.html)
    - [Reading VCF data](reading_vcf.html)
    - [Quality control](qc.html)
    - [Analysis of genomic data](analysis_of_genome.html)
    - [Analysis of GBS data](gbs_analysis.html)

- **About**
    - [About the authors](Authors.html)

- **Appendices**
    - [Data sets](Data_sets.html)
    - [Function glossary](funpendix.html)
    - [Source code for this primer](https://github.com/grunwaldlab/Population_Genetics_in_R/)
    - [Introduction to R](intro_to_R.html)