### 2021-11-04 WGSFullWalkthrough.Rmd

---
layout: default
title: WGS fastq2vcf walkthrough
parent: Whole Genome Resequencing
grand_parent: Choosing a PopGenomics Approach
editor_options: 
  chunk_output_type: console
---

## Walkthrough Data
In this walkthrough, data files from whole genome sequences at 15x coverage from Atlantic cod will be referenced. Their naming struction are a population ID in the form of "Pop#" with the number being which population the sample came from, a "\_"\, and a 5 digit sample ID which is two digits for the year it was sampled and a three digit sample ID number. Here is an example, "Pop6_18001". We will walk through steps from demultiplexing when unique indexes are used through variant calling with Freebayes.

# **Stacks**
If your samples need to be demultiplexed, running the program Stacks is a necessary first step. Stack’s process_radtags function searches sequence reads for barcodes that you provide and demultiplexes the reads using your barcode file. This file specifies which barcode belongs to which sample and will trim off those barcodes once it demultiplexes.

### Example scenario 
Samples used as examples throughout this walkthrough were multiplexed using 3 different unique sequences. Every sample had a unique combination of an i5 primer, i7 primer and adapter barcode. Data from the sequencing facility will demultiplex to the level of unique combinations of i5 and i7 primers. Therefore for these samples, there were 32 unique i5 and i7 primer combinations and so the sequencing facility sent 64 fastq.gz files: one for the forward sequence and one for the reverse for all 32 combinations. We will take one file as an example in the following code.

### Helpful Links
```process_radtags``` manual and specific walkthrough for ```process_radtags```  
[Manual](https://catchenlab.life.illinois.edu/stacks/manual/)  
[Walkthrough](https://catchenlab.life.illinois.edu/stacks/comp/process_radtags.php)

#### Example Fastq File from Sequencing Facility

##### Read 1
```
[schaal.s@login-00 CodGenomes]$ zcat i5-2-i7-9_R1_001.fastq.gz | head -n 10

@GWNJ-1012:218:GW191226406th:1:1101:1090:1000 1:N:0:GATCAG+ATAGAGGC
ACTTGACTGTGCGTTGGCCTGCGGGCTGACTCGGTCCTGAGATGGACTGCTGTGTAGTTTGAACCATAGATTCATTATATAGAACACGGTCTCCTCTGCGCTGCTGGCCAATGGAGCCGAACGTCCGCACTGGCGGGCGGCCATCTTGCC
+
FF:,FF,FFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFF:FFFFFFFFFFFFFFFFFFFFFFF,:FFFFFFFFF:FFFFFFF:FFFFFFFFFFFFF:FFFFFFF
@GWNJ-1012:218:GW191226406th:1:1101:1687:1000 1:N:0:GATCAG+ATAGAGGC
ACTTGATCCCTCTCACTCTCCTCTCCGTCTCCTCTTTTGTCCTCGTCTCTCTCCTCTCTCCCTCTCTCCCATCTCCCTCTCTATCAAGTAGATCGGAAGAGCACACGTCTGAACTCCAGTCACGATCAGATCTCGTATGCCGTCTTCTGC
+
FFFFFF,FFFFFFFFFFFFFF:FFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFF:FFFFFFFFFFFFFFFFFF:FFFFFFFFFFFFFFFF:FFFFFFFFFFFFFFFFFFFFFFFF:FFFF:::F,F::FFFFF
@GWNJ-1012:218:GW191226406th:1:1101:6840:1000 1:N:0:GATCAG+ATAGAGGC
ACTTGAAAAAAATACATAGCGGCCATGGACAGGATGACCTCTATGACAATGATAGAAACAGAAAGGACGCGGAGACTCTTGAGTCATCAAGTAGATCGGAAGAGCACACGTCTGAACTCCAGTCACGATCAGATCTCGTATGCCGTCTTC
```	
##### Read 2
```
[schaal.s@login-00 CodGenomes]$ zcat i5-2-i7-9_R2_001.fastq.gz | head -n 10

@GWNJ-1012:218:GW191226406th:1:1101:1090:1000 2:N:0:GATCAG+ATAGAGGC
ACTTGAAGGCAAGATGGCCGCCCGCCAGTGCGGACGTTCGGCTCCATTGGCCAGCAGCGCAGAGGAGACCGTGTTCTATATAATGAATCTATGGTTCAAACTACACAGCAGTCCATCTCAGGACCGAGTCAGCCCGCAGGCCAACGCACA
+
FFFFFF,FFFF:FFFFFF:FFFFFFFFFFFFFFFFFF:FFFFFFF:,FFFFF:FF,FFFFFFFFFFFFFFF,F,:FFFF:F,FFFFF:FF,FFFF,FFFFFF,FFF,FF:FFFFF:F:FFFF,FFF:F:F:FFFFFFFFFFF,:FFFFFF
@GWNJ-1012:218:GW191226406th:1:1101:1687:1000 2:N:0:GATCAG+ATAGAGGC
ACTTGAAAGAGAGGGAGATGGGAGAGAGGGAGAGAGGAGAGAGACGAGGACAAAAGAGGAGACGGAGAGGAGAGTGAGAGGGATCAAGTAGATCGGAAGAGCGTCGTGTAGGGAAAGAGTGTGCCTCTATGTGTAGATCTCGGTGGTCGC
+
FFFFFF,FFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFF,FFFF:FFFFFFFFFFFFFFFFFFFFFFFFFFFF:FF,FF:
@GWNJ-1012:218:GW191226406th:1:1101:6840:1000 2:N:0:GATCAG+ATAGAGGC
ACTTGAAGACTCAAGAGTCTCCGCGTCCTTTCTGTTTCTATCATTGTCATAGAGGTCATCCTGTCCATGGCCGCTATGTATTTTTATCAAGTAGATCGGAAGAGCGTCGTGTAGGGAAAGAGTGTGCCTCTATGTGTAGATCTCGGTGGT
```

### Stacks - process_radtags

Barcode File: For each set of paired end files that need to be demultiplexed, a barcode file needs to be created. For example, for files i5-4-i7-11_R1_001.fastq.gz and i5-4-i7-11_R2_001.fastq.gz its barcode file is as follows:

```
ATCACG	ATCACG	Pop6_18001
CGATGT	CGATGT	Pop7_18173
TTAGGC	TTAGGC	Pop8_18148
TGGCCA	TGGCCA	Pop9_18064
ACAGTG	ACAGTG	Pop3_16228
GCCAAT	GCCAAT	Pop4_17210
CAGATC	CAGATC	Pop6_18035
ACTTGA	ACTTGA	Pop6_18038
GATCAG	GATCAG	Pop8_18130
TAGCTT	TAGCTT	Pop8_18113
GGCTAC	GGCTAC	Pop7_18190
CTTGCA	CTTGCA	Pop1_17327
```

This file contains the 12 barcodes that have the i5_4 and i7_11 index and which sample that combination belongs to. This file is tab delimited and because I had inline barcodes, meaning they were the same on both ends of the read, I have the same barcode sequence in the first and second column. Then the sample ID is on the right.

Sample naming convention is important for downstream steps. The way the example samples are named are first a “Pop” identifier for what a priori population they belonged to (see Sample Notes file for example population IDs). Then a 5-digit identifier for the specific sample from that population where the first two digits represent the year the sample was collected and the last 3 digits represent the individual sample ID from that sampling year.

We provide a simple R script to create the barcode files for each primer index combination. See file: Stacks_FileCreate.R

#### Additional Flags

```-p``` path to your genome files. If you have paired data (forward and reverse reads), put your two fastq files in the same folder and only have those two files in it. I created folders for each index combination to keep everything organized and easy to run stacks on.

```-o``` path to where you want your output files to go. For subsequent steps, you'll want all of these demultiplexed files in the same folder. Create a folder like "Stacks_Out" where you write all your files to.

```-b``` path to your barcode file

```--inline_inline ``` The barcode option flag has 6 different options and you have to use the proper flag for your barcodes or indexes need to be demultiplexed. Because my samples have inline barcodes on either end of the sequence I need to use the --inline_inline flag.

```-P``` If paired end sequencing was used you need to add the -P flag to indicate this.

```--disable_rad_check``` this flag is needed if you are doing anything other than RAD-seq. Stacks defaults to requiring cutsite sequences that are used in RAD-seq, but for WGS you won't have cut sites and need to disable the rad check.

```-r``` this flag allows Stacks to rescue any barcodes that are off by one bp

#### Code for running Stacks
```
process_radtags -P -p {Path to genome files} -o .{Path for output} -b {barcode file} --inline_inline --disable_rad_check -r

```

#### Example code for running on a Cluster 
```
#!/bin/bash
#SBATCH --job-name=stacks_i54_i711                # sets the job name
#SBATCH --mem=1Gb                                 # reserves 10 GB memory
#SBATCH --partition=lotterhos                     # requests that the job is executed in lotterhos partition 
#SBATCH --mail-user=schaal.s@northeastern.edu
#SBATCH --mail-type=FAIL
#SBATCH --time=24:00:00                           # reserves machines/cores for 24 hours.
#SBATCH --output=stacksi54_i711.%j.out                # sets the standard output to be stored in file, where %j is the job id
#SBATCH --error=stacksi54_i711.%j.err                 # sets the standard error to be stored in file

process_radtags -P -p i54_i711/GenomeFilesi54_i711/ -o ../Stacks_Out/ -b i54_i711/barcodeFilei54_i711.txt --inline_inline --disable_rad_check -r
```

### Results 
Time: 				each file took approx. 7 hrs  
Memory: 			approx. 80 MB RAM

#### Read 1 Post Stacks for example sample Pop3_17304

```
[schaal.s@login-00 Stacks_Out]$ zcat Pop3_17304.1.fq.gz | head -n 10

@218_1_1101_1090_1000/1
CTGTGCGTTGGCCTGCGGGCTGACTCGGTCCTGAGATGGACTGCTGTGTAGTTTGAACCATAGATTCATTATATAGAACACGGTCTCCTCTGCGCTGCTGGCCAATGGAGCCGAACGTCCGCACTGGCGGGCGGCCATCTTGCC
+
,FFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFF:FFFFFFFFFFFFFFFFFFFFFFF,:FFFFFFFFF:FFFFFFF:FFFFFFFFFFFFF:FFFFFFF
@218_1_1101_1687_1000/1	TCCCTCTCACTCTCCTCTCCGTCTCCTCTTTTGTCCTCGTCTCTCTCCTCTCTCCCTCTCTCCCATCTCCCTCTCTATCAAGTAGATCGGAAGAGCACACGTCTGAACTCCAGTCACGATCAGATCTCGTATGCCGTCTTCTGC
+
,FFFFFFFFFFFFFF:FFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFF:FFFFFFFFFFFFFFFFFF:FFFFFFFFFFFFFFFF:FFFFFFFFFFFFFFFFFFFFFFFF:FFFF:::F,F::FFFFF
@218_1_1101_6840_1000/1
AAAAAATACATAGCGGCCATGGACAGGATGACCTCTATGACAATGATAGAAACAGAAAGGACGCGGAGACTCTTGAGTCATCAAGTAGATCGGAAGAGCACACGTCTGAACTCCAGTCACGATCAGATCTCGTATGCCGTCTTC
```

#### Read 2 Post Stacks for example sample Pop3_17304 
```
[schaal.s@login-00 Stacks_Out]$ zcat Pop3_17304.2.fq.gz | head -n 10

@218_1_1101_1090_1000/2
AGGCAAGATGGCCGCCCGCCAGTGCGGACGTTCGGCTCCATTGGCCAGCAGCGCAGAGGAGACCGTGTTCTATATAATGAATCTATGGTTCAAACTACACAGCAGTCCATCTCAGGACCGAGTCAGCCCGCAGGCCAACGCACA
+
,FFFF:FFFFFF:FFFFFFFFFFFFFFFFFF:FFFFFFF:,FFFFF:FF,FFFFFFFFFFFFFFF,F,:FFFF:F,FFFFF:FF,FFFF,FFFFFF,FFF,FF:FFFFF:F:FFFF,FFF:F:F:FFFFFFFFFFF,:FFFFFF
@218_1_1101_1687_1000/2
AAGAGAGGGAGATGGGAGAGAGGGAGAGAGGAGAGAGACGAGGACAAAAGAGGAGACGGAGAGGAGAGTGAGAGGGATCAAGTAGATCGGAAGAGCGTCGTGTAGGGAAAGAGTGTGCCTCTATGTGTAGATCTCGGTGGTCGC
+
,FFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFF,FFFF:FFFFFFFFFFFFFFFFFFFFFFFFFFFF:FF,FF:
@218_1_1101_6840_1000/2
AGACTCAAGAGTCTCCGCGTCCTTTCTGTTTCTATCATTGTCATAGAGGTCATCCTGTCCATGGCCGCTATGTATTTTTATCAAGTAGATCGGAAGAGCGTCGTGTAGGGAAAGAGTGTGCCTCTATGTGTAGATCTCGGTGGT
```

<br>
<br>

# **Quality Trimming**
Before sequence alignment to a reference genome, you need to trim your sequence data of bases that you have low confidence in being correct. This is to ensure that you have high confidence in both your alignment and the downstream variant calls. There are a few programs that do this including Trimmomatic and Fast P, but this walkthrough will describe Fast P and settings used for this program.

### Helpful Links  
[Fast P Manual](https://github.com/OpenGene/fastp#:~:text=fastp%20supports%20global%20trimming%2C%20which,or%20%2D%2Dtrim_tail1%3D1%20option.)

#### Parameters

```--in1``` This is read 1 for a given sample that uses paired end sequencing. For paired end sequencing, you will have two input files that need to be identified by ```--in1``` and ```--in2```. If doing single end sequencing, you only need the ```--in``` flag.

```--in2``` This is read 2 for a given sample that uses paired end sequencing.

```--out1``` This is the location and name of the out file for the read 1.

```--out2``` This is the location and name of the out file for the read 2.

```-q``` this is used as a filter on the PHRED quality score that each base is given from the sequencing facility. PHRED quality scores range from 0-40 with 40 being the highest quality. Here I removed any base that had a quality score of less than or equal to 15.

```-u``` this is used as a filter on the unqualified percent limit by specifying how many bases are allowed to be unqualified (0-100).

```--trim_front1``` This flag will trim the front base of read 1 by 1. This is necessary if you have an A-overhang on your reads that needs trimming.

```--trim_front2``` This flag will trim the front base of read 2 by 1.

```--cut_front``` this moves a sliding window, based on the size you set with the ```--cut_window_size```, flag from the front 5' to tail and drops bases in the window if its mean quality is below what you set as ```--cut_front_mean_quality```

```--cut_tail``` this moves a sliding window, based on the size you set with the ```--cut_window_size```, flag from the tail 3' end to front and drops bases in the window if its mean quality is below what you set as ```--cut_front_mean_quality```

```--cut_window_size``` the window size option for ```--cut_front```,  Range: 1-1000. The window can slide from either 5′ to 3′ or from 3′ to 5′, and the average quality score within the window is evaluated. If the average quality is lower than the threshold set in ```--cut_mean_quality``` then the bases in the window will be discarded

```--cut_mean_quality``` This is the mean quality requirement option shared by ```--cut_front```, ```--cut_tail``` or ```--cut_sliding```. Range: 1-36

```-j``` outputs a json format report to the file you specify

```-h``` outputs an html format report to the file you specify

#### Code for running fastp with paired end data
```
fastp --in1 {Path to fastq file for forward reads} --in2 {Path to fastq file for reverse reads} --out1 {Path to output quality trimmed forward reads} --out2 {Path to output quality trimmed forward reads} -q {min quality score to keep} -u {percent of unqualified bases allowed} --trim_front1 {number of bases to trim from front of forward read} --trim_front2 {number of bases to trim from front of reverse read} --cut_front --cut_tail --cut_window_size {window size to assess quality} --cut_mean_quality {mean min quality to cut} -j {path to output json file} -h {path to output html file} &> {path to output .log file}

```

#### Example for running on a Cluster
```
#!/bin/bash
#SBATCH --job-name=Pop9_18101_submitFastP.txt
#SBATCH --mem=2Gb
#SBATCH --mail-user=schaal.s@northeastern.edu
#SBATCH --mail-type=FAIL
#SBATCH --partition=lotterhos
#SBATCH --time=1:00:00
#SBATCH --nodes=1
#SBATCH --tasks-per-node=1
#SBATCH --output=FastP_Out/jobSum/clustOut/Pop9_18101.%j.out
#SBATCH --error=FastP_Out/jobSum/clustOut/Pop9_18101.%j.err

fastp --in1 Stacks_Out/Pop9_18101.1.fq.gz --in2 Stacks_Out/Pop9_18101.2.fq.gz --out1 FastP_Out/Pop9_18101.R1.fq.gz --out2 FastP_Out/Pop9_18101.R2.fq.gz -q 15 -u 50 --trim_front1 1 --trim_front2 1 --cut_front --cut_tail --cut_window_size 5 --cut_mean_quality 15 -j FastP_Out/jobSum/Pop9_18101.fp.json -h FastP_Out/jobSum/Pop9_18101.fp.html &> FastP_Out/jobSum/Pop9_18101.fp.trim.log
```

#### Results
Time: 	20-30 min per sample
Memory: 1.5 - 2GB of RAM

#### Other programs
Trimmomatic
Stacks

<br>
<br>

# **Alignment**
There are multiple programs used for aligning short reads to a reference: Bowtie, Bowtie2, BWA, STAR, etc. Here I will describe the steps I took using BWA.

### Helpful Links
[bwa Github](https://github.com/lh3/bwa)
[bwa Manual](http://bio-bwa.sourceforge.net/bwa.shtml)

### First step: index your genome 
Download your genome from the [NCBI database](https://www.ncbi.nlm.nih.gov/) or [UCSC database](http://hgdownload.cse.ucsc.edu/downloads.html) and run bwa index to index the genome. This allows BWA to essentially make a catalog (i.e., index) of the locations in the genome to make alignment faster.

#### Code for running BWA index
```
bwa index {Path to .fna genome file}
```

#### Example code for running on a cluster
```
#!/bin/bash
#SBATCH --job-name=subGenomesBWA                              
#SBATCH --mem=2Gb
#SBATCH --mail-user=schaal.s@northeastern.edu
#SBATCH --mail-type=FAIL
#SBATCH --partition=lotterhos
#SBATCH --time=6:00:00
#SBATCH --nodes=1
#SBATCH --tasks-per-node=1                        
#SBATCH --output=indexGenomeBWA.%j.out                
#SBATCH --error=indexGenomeBWA.%j.err

bwa index Cod_genome/GCF_902167405.1_gadMor3.0_genomic.fna
```
#### Results
Time: 9 minutes 
Memory: 1 GB of RAM

### Second step: alignment with BWA-MEM
There are a number of parameters to change in the bwa alignment. It is worth playing with parameters and outputting summary information to evaluate how well the alignment worked.

Dr. Jon Puritz provides an open sourced [guide](https://github.com/jpuritz/Winter.School2018/blob/master/Exercises/Day1/Mapping%20Exercise.md) for testing parameters to ensure you get the best alignment results (DOUBLE CHECK WITH HIM THAT IT IS OKAY FOR US TO PUT THIS IN HERE). General advice is that a gap penalty of 5 and a mismatch penalty of 3 consistently gives better results with a lot of marine species. However, it’s worth doing a couple of test runs with 3–5 samples comparing that to the default.

#### Arguments

```-B``` mismatch penalty used to adjust alignment scores based on the number of mismatches in the aligned sequence

```-O``` gap penalty used to adjust alignment scores based on the number and length of gaps and mismatches are when there are bases that deviate from the reference that the read is aligning to. This helps alleviate sequencing errors influencing the alignment.

```-a``` output all found alignments for SE or unpaired paired end; flagged as secondary alignments

```-M``` mark shorter split hits as secondary

```-R``` RGline (reference) because you will eventually be merging files this flag is important to add in the sample ID associated with these reads in the actual alignment file. This will remain in the file then for downstream analyses.

#### Code for running BWA index
```
bwa mem -O {num} -B {num} -a -M -R {Path to .fna genome file} {Path to sample .fg.gz forward read} {Path to sample .fg.gz reverse read}  > {Path to outfile location for .sam file}

```

#### Example for running on a Cluster

```
#!/bin/bash
#SBATCH --job-name=Pop3_17304_submitBWA
#SBATCH --mem=3Gb
#SBATCH --mail-user=schaal.s@northeastern.edu
#SBATCH --mail-type=FAIL
#SBATCH --partition=lotterhos
#SBATCH --time=15:00:00
#SBATCH --nodes=1
#SBATCH --tasks-per-node=1
#SBATCH --output=BWA_Out/clustOut/Pop3_17304.%j.out
#SBATCH --error=BWA_Out/clustOut/Pop3_17304.%j.err

bwa mem -O 5 -B 3 -a -M -R BWA_genome/GCF_902167405.1_gadMor3.0_genomic.fna FastP_Out/Pop3_17304.R1.fq.gz FastP_Out/Pop3_17304.R2.fq.gz  > BWA_Out/Pop3_17304aln.sam

```
#### Results (per sample):
Time: 7-10 hours 
Memory: about 1.5 GB

### Third Step: check alignment stats using samtools view and flagstat

#### Code for running samtools 
```
samtools view -Sbt {Path to .fna genome file} {Path to .sam aligned sample file} | samtools flagstat -
```

#### Example for running on a Cluster
```
#!/bin/bash
#SBATCH --job-name=Pop7_18156_alnCheckDef
#SBATCH --mem=2Gb
#SBATCH --mail-user=schaal.s@northeastern.edu
#SBATCH --mail-type=FAIL
#SBATCH --partition=lotterhos
#SBATCH --time=4:00:00
#SBATCH --nodes=1
#SBATCH --tasks-per-node=1
#SBATCH --output=clustOut/Pop7_18156_def.%j.out
#SBATCH --error=clustOut/Pop7_18156_def.%j.err

samtools view -Sbt ../BWA_genome/GCF_902167405.1_gadMor3.0_genomic.fna ../BWA_Out/Pop7_18156alnDef.sam | samtools flagstat -
```

#### example output from flagstat (sample: Pop1_16216):
```
75779849 + 0 in total (QC-passed reads + QC-failed reads)
0 + 0 secondary
1183259 + 0 supplementary
0 + 0 duplicates
74456805 + 0 mapped (98.25% : N/A)
74596590 + 0 paired in sequencing
37298295 + 0 read1
37298295 + 0 read2
70915282 + 0 properly paired (95.07% : N/A)
73157116 + 0 with itself and mate mapped
116430 + 0 singletons (0.16% : N/A)
2026064 + 0 with mate mapped to a different chr
739870 + 0 with mate mapped to a different chr (mapQ>=5)

```
Here you want to maximize your properly paired reads and percent mapped to the genome. Try a few parameters and see how you can maximize this.

Line 1 - total: the total number of alignments that pass the quality filter, this isn’t the total number  of reads. To get the total number of reads, subtract the secondary, supplementary, and duplicate values from this number which will be R1 + R2.  
Line 2 - secondary: (0x100 bit set)
Line 3 - supplementary: Supplementary alignments are those alignments that are part of a chimeric alignment (0x800 bit set)
Line 4 - duplicates: these mark potential PCR duplicates. This value should be zero if you filtered out all the pcr duplicates using picard tools (0x400 bit set)
Line 5 - mapped: (0x4 bit not set)
Line 6 - paired in sequencing: (0x1 bit set)
Line 7 - read1: (both 0x1 and 0x40 bits set)
Line 8 - read2: (both 0x1 and 0x80 bits set)
Line 9 - properly paired: (both 0x1 and 0x2 bits set and 0x4 bit not set)
Line 10 - with itself and mate mapped: (0x1 bit set and neither 0x4 nor 0x8 bits set)
Line 11 - singletons: Singleton is a mapped read whose mate is unmapped (both 0x1 and 0x8 bits set and bit 0x4 not set)
Line 12 - with mate mapped to different chr: low quality reads mapping to different chromosomes
Line 13 - with mate mapped to different chr (mapQ>=5): higher quality reads mapping to different chromosome

<br>
<br>

# **Sort and index your alignment**
Once you have aligned your reads to the reference you need to sort and index them for downstream use. This can be done using samtools.

### Helpful links
[Samtools Manual](http://www.htslib.org/doc/samtools.html)

### Step 1: convert your sam files to bam
Bam files are much smaller than sam files. By converting your files to Bam, you willincrease downstream efficiency.

####  Code for running samtools for Sam to Bam conversion
```
samtools view -S -b {Path to .sam file} > {Path for output .bam file}
```

#### Example for running on a Cluster
```
#!/bin/bash
#SBATCH --job-name=Pop4_17226_alnSamToBam
#SBATCH --mem=2Gb
#SBATCH --mail-user=schaal.s@northeastern.edu
#SBATCH --mail-type=FAIL
#SBATCH --partition=lotterhos
#SBATCH --time=4:00:00
#SBATCH --nodes=1
#SBATCH --tasks-per-node=1
#SBATCH --output=samtools_samToBam_Out/clustOut/Pop4_17226.%j.out
#SBATCH --error=samtools_samToBam_Out/clustOut/Pop4_17226.%j.err

samtools view -S -b BWA_Out/Pop4_17226aln.sam > samtools_samToBam_Out/Pop4_17226aln.bam
```
#### Results (per sample):
Time: 35-40 minutes 
Memory: about 20MB

<br>

### Step 2: Sort your bam files
Sorting files takes a lot of memory, but not a lot of time. There is one trick about the memory part to keep in mind. You need to set memory for the cluster, but you also need to set the memory allotment allowed for samtools sort to store data in. If you make the memory in samtools too low it will create 100s to 1000s of temporary files that it will later need to knit back together. For the example shown here, the memory was increased so that samtools didn't have to create temporary files. HOWEVER, you need to set the memory to a value that is lower than what you give to the cluster. If it is exactly the same or near the amount you gave to the cluster, it will fail and say you will get an error saying the program went over the memory limits. Therefore, for this example 40GB of memory was set in the samtools code and 50 GB of memory was set for each sample on the cluster.

#### Code for running samtools for sorting bam file
```
samtools sort -m {memory allocation to samtools} {Path to .bam file} > {Path for output sorted .bam file}
```

#### Example for running on a Cluster
```
#!/bin/bash
#SBATCH --job-name=Pop4_17226_alnSortedBam
#SBATCH --mem=50Gb
#SBATCH --mail-user=schaal.s@northeastern.edu
#SBATCH --mail-type=FAIL
#SBATCH --partition=lotterhos
#SBATCH --time=24:00:00
#SBATCH --nodes=1
#SBATCH --tasks-per-node=1
#SBATCH --output=samtools_sortedBam_Out/clustOut/Pop4_17226.%j.out
#SBATCH --error=samtools_sortedBam_Out/clustOut/Pop4_17226.%j.err

samtools sort -m 40G samtools_samToBam_Out/Pop4_17226aln.bam > samtools_sortedBam_Out/Pop4_17226aln.sorted.bam
```

#### Results (per sample):
Time: about 10 min  
Memory: 40GB

<br>

### Step 3: Index your sorted bam file

#### Code for running samtools for indexing sorted bam file
```
samtools index {Path to sorted .bam file} > {Path for output sorted and indexed .bam file}
```

#### Example for running on a Cluster
```
#!/bin/bash
#SBATCH --job-name=Pop4_17226_alnIndexBam
#SBATCH --mem=50Mb
#SBATCH --mail-user=schaal.s@northeastern.edu
#SBATCH --mail-type=FAIL
#SBATCH --partition=lotterhos
#SBATCH --time=2:00:00
#SBATCH --nodes=1
#SBATCH --tasks-per-node=1
#SBATCH --output=samtools_indexedBam_Out/clustOut/Pop4_17226.%j.out
#SBATCH --error=samtools_indexedBam_Out/clustOut/Pop4_17226.%j.err

samtools index samtools_sortedBam_Out/Pop4_17226aln.sorted.bam > samtools_indexedBam_Out/Pop4_17226aln_sorted_indexed
```

#### Results (per sample):
Time: about 3-5 min  
Memory: about 15 MB

<br>
<br>

# **Checking alignment**

### First step: identifying regions to calculate coverage
It is important to check your coverage along each chromosome to determine whether you have regions of low or no coverage that could bias downstream analyses. One way to evaluate this is to first download the GFF file from NCBI for the genome of your species. As an example here is a link to the Atlantic cod genome used for the analysis examples in this walkthrough: [Atlantic cod genome](https://www.ncbi.nlm.nih.gov/genome/?term=txid8049[orgn])

Then subset the genome file for the chromosomes or genomic regions you want to evaluate coverage. Below is an example of subsetting the Atlantic cod genome for all 23 chromosomes in the R programming language: 
```
cod.gff <- read.table("src/alignment/GCF_902167405.1_gadMor3.0_genomic.gff", sep = "\t", quote = "")
scaffolds.gff <- cod.gff[cod.gff$V3 == "region",]
chrom.gff <- scaffolds.gff[1:23,]
write.table(chrom.gff, "src/alignment/GCF_902167405.1_gadMor3.0_genomic_chroms.gff", row.names = FALSE, 
            sep = "\t", col.names = FALSE, quote = FALSE)
```

### Second step: calculating coverage

#### Code for running bedtools coverage 
```
bedtools coverage -a {Path to GFF file created in previous step} -b {Path to aligned, sorted and indexed .bam file} -sorted -d > {Path for output .txt file}

```

#### Example for running on a Cluster
```
#!/bin/bash
#SBATCH --job-name=Pop1_16216_bedCov
#SBATCH --mem=50Gb
#SBATCH --mail-user=schaal.s@northeastern.edu
#SBATCH --mail-type=FAIL
#SBATCH --partition=lotterhos
#SBATCH --time=4:00:00
#SBATCH --nodes=1
#SBATCH --tasks-per-node=1
#SBATCH --output=bedtools_coverage/clustOut/Pop1_16216bedCov.%j.out
#SBATCH --error=bedtools_coverage/clustOut/Pop1_16216bedCov.%j.err
bedtools coverage -a Cod_genome_data/GCF_902167405.1_gadMor3.0_genomic_chroms.gff -b samtools_sortedBam_Out/Pop1_16216aln.sorted.bam -sorted -d > bedtools_coverage/Pop1_16216.coverageCalcDflag.txt 
```
```-a``` reference genome

```-b``` sorted bam file

```-d``` give per base coverage

```-sorted``` tells bedtools that this file is already sorted

### Third step: subsetting
Subset for just the chromosome data and the columns of interest in the output file (reduces file sizes from 111GB to 17GB this part can easily be piped so you aren't creating that large intermediate file)  
```
#!/bin/bash
#SBATCH --job-name=Pop1_16216_alnCheck
#SBATCH --mem=2Gb
#SBATCH --mail-user=schaal.s@northeastern.edu
#SBATCH --mail-type=FAIL
#SBATCH --partition=lotterhos
#SBATCH --time=4:00:00
#SBATCH --nodes=1
#SBATCH --tasks-per-node=1
#SBATCH --output=bedtools_coverage/clustOut/Pop1_16216awk.%j.out
#SBATCH --error=bedtools_coverage/clustOut/Pop1_16216awk.%j.err
awk -F"\t" '$1~/NC*/' bedtools_coverage/Pop1_16216.coverageCalcDflag.txt | awk '{print $1,$11,$12}' > bedtools_coverage/Pop1_16216.coverageCalcChr.txt
```

### Fourth step: calculate coverage average 
run the output file through the R script for calculating averages for different window sizes and plot results in ggplot   
```

### PROCESS COVERAGE DATA TO VISUALIZE ACROSS THE GENOME ###
## Sara M. Schaal

## LOAD LIBRARIES
library(ggplot2)
library(data.table)

## USER INPUTS
setwd("/scratch/schaal.s/CodGenomes") # SET YOUR WORKING DIRECTORY
samps <- c("Pop1_16216", "Pop1_17291", "Pop4_17236", "Pop5_17278", "Pop6_18017") # INPUT SAMPLES OF INTEREST
colors <- c("steelblue2", "chartreuse3", "orchid2", "firebrick1", "goldenrod2") # INPUT COLORS 
n <- 10000 # INPUT SIZE OF WINDOWS TO CALCULATE AVERAGE COVERAGE

## PROCESS DATA 
# first step through each sample
for(i in 1:length(samps)){
  df <- fread(paste0("bedtools_coverage/", samps[i], ".coverageCalcChr.txt"), 
                                                      sep =" ", quote = "", data.table = FALSE)
  colnames(df) <- c("chrom", "base", "coverage")
  chroms <- unique(df[,1])
  df.sample.data <- NULL
  # step through each chromosome and break into the increment chunks you set with n
  for(j in 1:length(chroms)){
    df.chrom <- df[df$chrom == chroms[j], ]
    extra <- nrow(df.chrom) %% n
    # this next part is taking our windows set by n and giving each window a number id
    # this grouped dataframe is made by binding the original df.chrom with grouping values for the window size you want 
    # then finding what the last grouping value would be using because it will most like not be an even increment of your n
    # for example if you divide a chromosome by your n and get 3000.3 then you can easily input the first 3000
    # grouping values in with rep (middle part of the following cbind function) then the last grouping value will be 3001 which
    # you get by rounding using ceiling in the last part of this cbind
    grouped <- cbind(df.chrom, c(rep(1:(nrow(df.chrom)/n), each = n), rep(ceiling(nrow(df.chrom)/n), extra)))
    colnames(grouped)[4] <- "grouping"
    # now bind this chromosomes data in the full dataframe
    df.sample.data <- rbind(df.sample.data, grouped)
  }
  # finally take your new dataframe and calculate the average coverage for your increments using
  # this new grouping variable and the chromosome
  df.covAve <- aggregate(coverage~grouping + chrom, data = df.sample.data, FUN = mean)

  #### PLOTTING ####
  pdf(paste0("figures/", samps[i], "MaxcoveragePlot", n, ".pdf"), height= 15, width=15)

  ## I make two plots because there will undoubtedly be some loci that have really high coverage which makes
  # it hard to see all the spread of the majority of the data. This first graph is set to the max coverage
  # found in the data frame and the second plot is setting your y limit to a more reasonable value for your
  # data. For me 100 x was good but feel free to change to what is appropriate for your data.

  print(ggplot(data = df.covAve, aes(x = grouping, y = coverage)) +
    geom_point(col = colors[i], alpha = 0.5) +
    facet_wrap(~chrom) +
    labs(y = "Coverage", x = paste0("location every ", n, " bases"), 
         title = paste0(samps[i], "Genome Coverage up to Max Coverage")) +
    ylim(0, max(df.covAve$coverage)) +
    xlim(0, max(df.covAve$grouping)) +
    theme_bw() + 
    theme(panel.border = element_blank(), panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          axis.line = element_line(colour = "black"), 
          legend.position = "none"))
  dev.off()

  pdf(paste0("figures/", samps[i], "100XcoveragePlot", n, ".pdf"), height= 15, width=15)

  print(ggplot(data = df.covAve, aes(x = grouping, y = coverage)) +
          geom_point(col = colors[i], alpha = 0.5) +
          facet_wrap(~chrom) +
          labs(y = "Coverage", x = paste0("location every ", n, " bases"),
               title = paste0(samps[i], "Genome Coverage up to 100X")) +
          ylim(0, 100) +
          xlim(0, max(df.covAve$grouping)) +
          theme_bw() + 
          theme(panel.border = element_blank(), panel.grid.major = element_blank(), 
                panel.grid.minor = element_blank(),
                axis.line = element_line(colour = "black"), 
                legend.position = "none"))
  dev.off()

}
```
<!-- <img src="./Pop1_17291100XcoveragePlot10000.png" width="500">  -->

#### Results (R script for 5 samples):
Time: about 4.5 hours  
Memory: about 85 GB

<br>
<br>

# **Picard MarkDuplicates**

The tool's main output is a new SAM or BAM file, in which duplicates have been identified in the SAM flags field for each read. Duplicates are marked with the hexadecimal value of 0x400, which corresponds to a decimal value of 1024. For the example data shown in this walkthrough,  OPTICAL_DUPLICATE_PIXEL_DISTANCE=2500 needs to be set which is for data coming from a patterned flow cell. The example whole genome data comes from a NovaSeq run which uses the patterned flow cell. The default for that flag is only 100 which is appropriate for unpatterned flow cells. Make sure you know which technology your data come from and set this parameter as needed.

### Helpful Links
[Picard MarkDuplicates Manual](https://gatk.broadinstitute.org/hc/en-us/articles/360037052812-MarkDuplicates-Picard-)
[Breakdown and explanation of different duplicates from Illumina sequencing](http://core-genomics.blogspot.com/2016/05/increased-read-duplication-on-patterned.html)

#### Code for running PicardTools
```
picard MarkDuplicates I={Path to aligned, sorted and indexed .bam file} O={Path for output of duplicated marked .bam file} M={Path for metrics output file} OPTICAL_DUPLICATE_PIXEL_DISTANCE={Value specific to sequencing technology} TAGGING_POLICY={Type of duplicates to mark} &> {Path to .log file}

```

#### Example for running on a Cluster
```
#!/bin/bash
#SBATCH --nodes=1
#SBATCH --time=8:00:00
#SBATCH --job-name=picard_benchmark
#SBATCH --partition=short
#SBATCH --cpus-per-task=8
#SBATCH --output=picard_log.o
#SBATCH --error=picard_log.e

picard MarkDuplicates I=../samtools_sortedBam_Out/Pop9_18093aln.sorted.bam O=../picard_Out/Pop9_18093aln.sorted.md.bam M=md_metrics.txt OPTICAL_DUPLICATE_PIXEL_DISTANCE=2500 TAGGING_POLICY=OpticalOnly &> pop9_18093aln.sorted.md.log
```
#### Results (per sample):
Time: about 30 min  
Memory: about 2-3GB

<br>
<br>

# **Samtools filtering**

Before running the SNP caller, you need to filter the samples to ensure you are only using good quality aligned data. To do this, the bam file is filtered using samtools view.

#### Code from dDocent for quality filtering: 
```
samtools view -@32 -h -q 10 -F 0x100 -F 0x400 $1-RGmd.bam | mawk '$6 !~/[8-9].[SH]/ && $6 !~ /[1-9][0-9].[SH]/' | samtools view -@ 32 -b
```

In dDocent, all reads are filtered if mapping quality is less than 10, removes PCR duplicates (marked in previous Picard step this quality filter then removes them) with the -F 0x400 flag and removes non primary alignments with the 0x100. The mawk line of code then removes hard and soft clipped reads. Finally, the last samtools view code exports the results as a bam file.

```-@``` option allocates threads for the job

```-h``` include the header in the output

```-q``` skip alignments with MAPQ smaller than INT[0]

```-F``` Do not output alignments with any bits set in INT present in the FLAG field. INT can be specified in hex by beginning with 0x' (i.e. /^0x[0-9A-F]+/) or in octal by beginning with 0' (i.e. /^0[0-7]+/) [0]

```-b``` output in the bam format

#### Example for an altered dDocent code
Changed the mawk code to not remove the hard clipped reads because I am interested in inversion breakpoints. This would most likely filter those out.

#### Example for running on a cluster
```
#!/bin/bash
#SBATCH --job-name=samtoolsFilter
#SBATCH --mem=750Mb
#SBATCH --mail-user=schaal.s@northeastern.edu
#SBATCH --mail-type=FAIL
#SBATCH --partition=lotterhos
#SBATCH --time=0:30:00
#SBATCH --cpus-per-task=2
#SBATCH --output=../samtools_filter_Out/clustOut/samtoolsFilter.%j.out
#SBATCH --error=../samtools_filter_Out/clustOut/samtoolsFilter.%j.err

samtools view -@2 -h -q 10 -F 0x100 -F 0x400 ../picard_Out/Pop1_16216aln.sorted.md.bam | mawk '$6 !~ /[1-9][0-9].[SH]/'| samtools view -@2 -b > ../samtools_filter_Out/Pop1_16216.f.bam

```
#### Results (per sample):
Time: 15-20 min  
Memory: about 500MB

<br>
<br>

# **Merge Reads**

Before calling SNPs you need to merge all your per sample Bam files. This is why we had to add the RG ids during the mapping step. This ensures all reads have a unique identifier for which sample they belong to once merged. By merging all the reads, the SNP caller doesn’t need to keep opening and closing individual sample files, which will increase the runtime. You need to make a text file that includes all the bamfile names that you want to merge and then run the line of code below. Because this is computationally intensive, I used 32 cpus on a single compute node.

### Helpful Links
[Samtools Merge Manual](http://www.htslib.org/doc/samtools-merge.html)

#### Code for running samtools merge
```
BAMLIST={Path to .txt Bamfile list}
samtools merge {Path for output of merged .bam file} -b ${BAMLIST} -@{number of threads to use}

```

#### Example for running on a Cluster
```
#!/bin/bash
#SBATCH -p short
#SBATCH --nodes 1
#SBATCH --cpus-per-task=32
#SBATCH -t 120:00:00
#SBATCH --constraint=zen2
#SBATCH --mem=5G
#SBATCH -o clustOut/slurm.%N.%j.out
#SBATCH -e clustOut/slurm.%N.%j.err
#SBATCH --job-name="Samtools-merge_test"
#SBATCH --mail-type=END,FAIL
#SBATCH --mail-user=schaal.s@northeastern.edu

BAMLIST=./bamlist.txt
samtools merge mergedBam_n32_all.bam -b ${BAMLIST} -@32

```
#### Results (merging 296 samples):
Time: about 36 hours  
Memory: about 150MB

<br>
<br>

# **SNP calling**

There are two major variant callers that are widely used: GATK and Freebayes. Here we walk through using freebayes-parallel to optimize runtime on our cluster. SNP calling is computationally intensive so this step will take some work to optimize both the size of genome chunks you want to analyze at a time and how many cpus to run your samples on.

There are a few files that you need for this step: 1) your population list file which has a line per sample where the sample name is the first item in the line and the population ID is the second item in the line (see example header of a population list file below) 2) your .fna genome reference that you are calling snps from and 3) regions file where you tell freebayes what size chunks of the genome you want to analyze at a time. Depending on your coverage, this last file is very important to troubleshoot to maximize memory use vs. run times. In the example data, we found we were reaching about 90% efficiency on our cluster with 9-13 hour runs using 64 cores and allotting 90 GB of memory when we split the genome up into 100kb chunks. So this is how we proceeded and ran each chromosome individually by splitting them each up into 100kb chunks.

### Helpful links
[Freebayes Github Page](https://github.com/freebayes/freebayes)
[Freebayes README](https://ngseasy.readthedocs.io/en/latest/containerized/ngseasy_dockerfiles/ngseasy_freebayes/README/)

#### example header of a population list file
```
Pop4_17219	Pop4
Pop2_17008	Pop2
Pop9_18076	Pop9
Pop1_16232	Pop1
Pop7_18161	Pop7
Pop8_18147	Pop8
```
#### example header of a regions list file
```
NC_044048.1:1-100000
NC_044048.1:100001-200000
NC_044048.1:200001-300000
NC_044048.1:300001-400000
NC_044048.1:400001-500000

```

#### Code for running freebayes parallel
```
POPFILE={Path to population .txt file}
REF={Path to .fna reference genome file}
freebayes-parallel {path to regions .txt file} {num threads} -f ${REF} -b {Path to merged bam file} --populations ${POPFILE} -m {num} -q {num} -E {num} --min-repeat-entropy {num} -n {num} -F {fraction} >> {Path for .vcf out file}

```

#### Arguments

```-m```--min-mapping-quality Q. Exclude alignments from analysis if they have a mapping quality less than Q

```-q``` --min-base-quality Q. Exclude alleles from analysis if their supporting base quality is less than Q

```-E``` Allow complex alleles with contiguous embedded matches of up to this length

```-n``` --use-best-n-alleles N. Evaluate only the best N SNP alleles, ranked by sum of supporting quality scores.

```-F``` --min-alternate-fraction N. Require at least this fraction of observations supporting an alternate allele within a single individual in order to evaluate the position

#### Example for running on a Cluster
```
#!/bin/bash
#SBATCH -p long
#SBATCH --nodes 1
#SBATCH --cpus-per-task=64
#SBATCH -t 120:00:00
#SBATCH --constraint=zen2
#SBATCH --mem=90GB
#SBATCH -o noRegion.%N.%j.out
#SBATCH -e noRegion.%N.%j.err
#SBATCH --job-name="Freebayes_parallel_test"
#SBATCH --mail-type=END,FAIL
#SBATCH --mail-user=schaal.s@northeastern.edu

POPFILE=/scratch/schaal.s/CodGenomes/10_freebayes/poplist.txt
REF=/scratch/schaal.s/CodGenomes/Cod_genome_data/GCF_902167405.1_gadMor3.0_genomic.fna
freebayes-parallel regionsFiles/NC_044048.1_100kbRegions.txt 128 -f ${REF} -b ../labeled_bam_Out/mergedBam_n128_all_lot.bam --populations ${POPFILE} -m 5 -q 5 -E 3 --min-repeat-entropy 1 -n 10 -F 0.1 >> outFiles/VarCall_freebayes-par.chrom_NC_044048.1.vcf

```

#### Results (running the 23 chromosomes across nodes with 64 cores each): 
Time: 6-12 hours each (total about 36 hours)   
Memory: 60-65GB

```{r, echo=FALSE,eval=FALSE}
library(knitr)
my.jekyll.site <- "marineomics.github.io"
KnitPost <- function(input, base.url = my.jekyll.site) {
  opts_knit$set(base.url = base.url)
  fig.path <- paste0("images/", sub(".Rmd$", "", basename(input)), "/")
  opts_chunk$set(fig.path = fig.path)
  opts_chunk$set(fig.cap = "center")
  render_jekyll()
  knit(input, envir = parent.frame())
}

setwd("~/Postdoc_Projects/RCN_MarineOmics/Website/marineomics.github.io/guidelines")
KnitPost("~/Postdoc_Projects/RCN_MarineOmics/Website/marineomics.github.io/guidelines/WGSFullWalkthrough.Rmd")

```