### 2021-11-04 dapc.Rmd

```{r,message = FALSE,echo = FALSE}
  library("knitcitations")
  library("knitr")
  cite_options(citation_format = "pandoc", max.names = 3, style = "html", hyperlink = "to.doc")
  bib <- read.bibtex("bibtexlib.bib")
  opts_chunk$set(tidy = FALSE, message = FALSE, warning = FALSE,
                 fig.width = 10, fig.height = 6, cache = TRUE)
  # use this to set knitr options:
  # http://yihui.name/knitr/options #chunk_options
```
---
title: 'Discriminant analysis of principal components (DAPC)'
subtitle: '*NJ Gr&uuml;nwald, ZN Kamvar, and SE Everhart*'
---

# Introduction
```{r}
library(ggplot2)
library(SeqArray)
library(tidyverse)
library(SNPRelate)
```

## Code chunk 
```{r}
x <- c(1,2,3,4)
y <- c(50, 99, 21, 20)
plot(x,y)
```

# References