### 2023-02-16 DNA-methylation.Rmd
taking through bismark

---
title: 'DNA Methylation Analysis'
subtitle: '*Steven Roberts* and *Sam White* '
bibliography: DNA-Methylation.bib
---

# Introduction
The purpose of this tutorial is to provide an example of taking Bisulfite treated DNA sequence data through identification of methylation at the loci and gene level. The principle behind creating bisulfite treated DNA libraries is that when cytosines are not methylated they will be converted to uracil, which ultimately result in thymine. When aligning to a reference genome the percent methylation at a given CpG loci is determined by examine the ratio to thymines and cytosines. For example if there is a given CpG loci with 30% of the alignment with the cytosine on the reference being and thymine, one would consider this cytosine loci (CpG) to be 70% methylated.

# Sequence Quality
An example of what raw data might look like from a marine invertebrate is shown. _Note that DNA methylation patters are diverse across taxa with most marine invertebrates demonstrating a mosiac pattern and global methylation at CpG loci is approximately 15%._

Following trimming one would expect to see something similar to this. This particular data set was trimmed with the following parameters:

```
/gscratch/srlab/programs/TrimGalore-0.4.5/trim_galore \
--output_dir /gscratch/scrubbed/strigg/analyses/20200311/WGBS_MBD \
--paired \
--fastqc_args \
--outdir /gscratch/scrubbed/strigg/analyses/20200311/WGBS_MBD/FASTQC \
--threads 28 \
--illumina \
--clip_R1 10 \
--clip_R2 10 \
--three_prime_clip_R1 10 \
--three_prime_clip_R2 10 \
--path_to_cutadapt /gscratch/srlab/programs/miniconda3/bin/cutadapt \
/gscratch/scrubbed/sr320/froger-raw/00_fastq/Meth17_R1_001.fastq.gz \
```

![bp-chart](images/meth-bp-chart.png)

# Read Alignment
The software used as part of this tutorial is **Bismark Bisulfite Mapper**. As with any software it is best to be famililar with the [User Manual](https://rawgit.com/FelixKrueger/Bismark/master/Docs/Bismark_User_Guide.html). Thus content below will provided with the assumption that the reader has read the manual and is meant to serve as guidance based on experience working with marine invertebrates.

The first step in the process in preparing the genome. Example code:

```
# Directories and programs
bismark_dir="/programs/Bismark-0.21.0"
bowtie2_dir="/programs/bowtie2-2.3.4.1-linux-x86_64/"
genome_folder="/where/the/fastafile/lives/"

${bismark_dir}/bismark_genome_preparation \
--verbose \
--parallel 28 \
--path_to_aligner ${bowtie2_dir} \
${genome_folder}

```

You should expect to prepared genome with new directories indluding

```
./Bisulfite_Genome
./Bisulfite_Genome/GA_conversion
./Bisulfite_Genome/CT_conversion
```

For doing the alignment of the trim reads to the genome the following code structured is used

```
find ${reads_dir}*_R1_001_val_1.fq.gz \
| xargs basename -s _R1_001_val_1.fq.gz | xargs -I{} ${bismark_dir}/bismark \
--path_to_bowtie ${bowtie2_dir} \
-genome ${genome_folder} \
-p 4 \
-score_min L,0,-0.6 \
--non_directional \
-1 ${reads_dir}{}_R1_001_val_1.fq.gz \
-2 ${reads_dir}{}_R2_001_val_2.fq.gz \
-o Mcap_tg
```

This will create bam files (sequence alignment files)

Files are then deduplicated if they are whole genome bisulfite treated samples.
This command will deduplicate the Bismark alignment BAM file and remove all reads but one which align to the the very same position and in the same orientation. This step is recommended for whole-genome bisulfite samples, but should not be used for reduced representation libraries such as RRBS, amplicon or target enrichment libraries.

```
find *.bam | \
xargs basename -s .bam | \
xargs -I{} ${bismark_dir}/deduplicate_bismark \
--bam \
--paired \
{}.bam
```

Methylation levels are then extracted using the `bismark_methylation_extractor`. For example:

```
${bismark_dir}/bismark_methylation_extractor \
--bedGraph \
--counts \
--comprehensive \
--merge_non_CpG \
--multicore 28 \
--buffer_size 75% \
*deduplicated.bam
```

This will create deduplicated.bismark.cov.gz, uncompressed is in this format.

```
NC_035784.1 141 141 37.5    3   5
NC_035784.1 142 142 100 2   0
NC_035784.1 155 155 70  7   3
NC_035784.1 156 156 100 2   0
NC_035784.1 291 291 0   0   2
NC_035784.1 292 292 0   0   3
NC_035784.1 313 313 0   0   1
NC_035784.1 314 314 66.6666666666667    2   1
NC_035784.1 470 470 66.6666666666667    4   2
NC_035784.1 611 611 0   0   4
```

```
<chromosome> <start position> <end position> <methylation percentage> <count methylated> <count unmethylated>
```

A genome-wide cytosine report ins generated, with this code as an example:

```
find *deduplicated.bismark.cov.gz \
| xargs basename -s _trimmed_bismark_bt2.deduplicated.bismark.cov.gz \
| xargs -I{} ${bismark_dir}/coverage2cytosine \
--genome_folder ${genome_folder} \
-o {} \
--merge_CpG \
--zero_based \
{}_trimmed_bismark_bt2.deduplicated.bismark.cov.gz
```

Generating  a file `.CpG_report.merged_CpG_evidence.cov`

```
NC_035785.1 217 219 100.000000  17  0
NC_035785.1 523 525 87.500000   7   1
NC_035785.1 556 558 50.000000   5   5
NC_035785.1 727 729 100.000000  16  0
NC_035785.1 1330    1332    0.000000    0   2
NC_035785.1 1403    1405    0.000000    0   2
NC_035785.1 1494    1496    66.666667   2   1
NC_035785.1 1747    1749    100.000000  8   0
NC_035785.1 2024    2026    100.000000  24  0
NC_035785.1 2054    2056    93.333333   14  1
```

From here it could be useful to do some simple file reorganizaion to obtain bedgraph or tab files for down stream analysis.

```

#creating bedgraphs post merge

for f in *merged_CpG_evidence.cov
do
  STEM=$(basename "${f}" .CpG_report.merged_CpG_evidence.cov)
  cat "${f}" | awk -F $'\t' 'BEGIN {OFS = FS} {if ($5+$6 >= 10) {print $1, $2, $3, $4}}' \
  > "${STEM}"_10x.bedgraph
done

for f in *merged_CpG_evidence.cov
do
  STEM=$(basename "${f}" .CpG_report.merged_CpG_evidence.cov)
  cat "${f}" | awk -F $'\t' 'BEGIN {OFS = FS} {if ($5+$6 >= 5) {print $1, $2, $3, $4}}' \
  > "${STEM}"_5x.bedgraph
done

#creating tab files with raw count for glms

for f in *merged_CpG_evidence.cov
do
  STEM=$(basename "${f}" .CpG_report.merged_CpG_evidence.cov)
  cat "${f}" | awk -F $'\t' 'BEGIN {OFS = FS} {if ($5+$6 >= 10) {print $1, $2, $3, $4, $5, $6}}' \
  > "${STEM}"_10x.tab
done

for f in *merged_CpG_evidence.cov
do
  STEM=$(basename "${f}" .CpG_report.merged_CpG_evidence.cov)
  cat "${f}" | awk -F $'\t' 'BEGIN {OFS = FS} {if ($5+$6 >= 5) {print $1, $2, $3, $4, $5, $6}}' \
  > "${STEM}"_5x.tab
done
```