### 2023-12-13 DRAFT_DNA_methylation.Rmd
Update DRAFT_DNA_methylation.Rmd

```{r,message = FALSE,echo = FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
library(knitcitations)
library(kableExtra)
opts_chunk$set(fig.width = 10,
               fig.height = 5,
               cache = FALSE)
cite_options(citation_format = "pandoc", max.names = 3, style = "html", 
             hyperlink = "to.doc")
```

---
title: 'DNA Methylation Analyses'
subtitle: '*Steven Roberts*, *Sam White*, and *Yaamini Venkataraman*'
bibliography: DRAFT_DNA_methylation/DNA-Methylation.bib
---
Initial publication year: TBD <br>
[How to cite](https://marineomics.github.io/#How_to_Cite)

# Introduction

Thanks to advances in sample preparation and sequencing methods, there is an influx of research examining epigenomics in non-model systems. Broadly, epigenetics can be defined as changes to gene expression that do not arise from changes in the DNA sequence. DNA methylation, or addition of a methyl (CH<sub>3</sub>) group to a cytosine base adjacent to a guanine (CpG) is one of the more commonly studies epigenetic mechanisms, partly because of its environmental sensitivity and potential role in phenotypic plasticity. For a good review of epigenetic mechanisms including DNA methylation in marine organisms see @doi:10.1146/annurev-marine-010318-095114.

The purpose of this tutorial is to provide an example of aligning bisulfite-treated and enzymatically converted DNA sequence data and discuss needs for potential downstream applications. The principle behind creating bisulfite-treated DNA libraries is that when cytosines are not methylated, then they will be converted to uracil, which ultimately result in thymine nucleotides in the sequence data. @https://doi.org/10.1111/1755-0998.13542 offers a comparison of three methods for quantifying DNA methylation at single base-pair resolution using whole genome bisulfite sequencing (WGBS), reduced representation bisulfite sequencing (RRBS), and methyl-CpG binding domain bisulfite sequencing (MBDBS). When aligning bisulfite-converted data to a reference genome the percent methylation at a given CpG loci is determined by examining the ratio of thymines to cytosines. For example if there is a given cytosine adjacent to a guanine (ie CpG locus) with 30% of the reads containing thymine, one would consider this cytosine loci (CpG) to be 70% methylated. Since bisulfite treatment can damage DNA, newer library preparation and sequencing methods (ex. [NEBNext® Enzymatic Methyl-seq Kit; EM-seq](https://www.neb.com/en-us/products/e7120-nebnext-enzymatic-methyl-seq-kit#Protocols,%20Manuals%20&%20Usage) are starting to gain popularity. As EM-Seq produces loci-level data, alignment of EM-Seq data is similar to WGBS.

The tutorial will cover the following:

- Sequence quality assessment
- Read alignment
- Methylation quantification
- File conversions
- Potential downstream applications. Although this tutorial will not provide code for downstream applications, it will link to other open-access resources and examples.

The tutorial below is based on WGBS from a species of marine intertebrate (*Montipora capitata* coral) examined in @https://doi.org/10.1111/1755-0998.13542, and all raw data can be accessed under NCBI Bioproject PRJNA691891 if you would like to follow the tutorial with this dataset. The general workflow would be consistent for other library preparation and sequencing methods that produce loci-level data. The tutorial indicates places where special analytical consideration should be given to other data types.

# Sequence Quality

An example of what raw data might look like from a marine invertebrate is shown. *Note that DNA methylation patters are diverse across taxa with most marine invertebrates demonstrating a mosiac pattern and global methylation at CpG loci is approximately 15%.*

Following trimming one would expect to see something similar to this. This particular data set was trimmed with the following parameters:

    /gscratch/srlab/programs/TrimGalore-0.4.5/trim_galore \
    --output_dir /gscratch/scrubbed/strigg/analyses/20200311/WGBS_MBD \
    --paired \
    --fastqc_args \
    --outdir /gscratch/scrubbed/strigg/analyses/20200311/WGBS_MBD/FASTQC \
    --threads 28 \
    --illumina \
    --clip_R1 10 \
    --clip_R2 10 \
    --three_prime_clip_R1 10 \
    --three_prime_clip_R2 10 \
    --path_to_cutadapt /gscratch/srlab/programs/miniconda3/bin/cutadapt \
    /gscratch/scrubbed/sr320/froger-raw/00_fastq/Meth17_R1_001.fastq.gz \

```{r meth-bp, echo = FALSE, out.width = "70%", fig.align = "center"}
knitr::include_graphics("DRAFT_DNA_methylation/meth-bp-chart.png")
```

For WGBS and MBDBS data, it is general practice to hard-trim 10 bp from the beginning and end of each read. [RRBS data often needs different trimming specifications](https://rawgit.com/FelixKrueger/Bismark/master/Docs/Bismark_User_Guide.html#ix-notes-about-different-library-types-and-commercial-kits).

# Read Alignment

The software used as part of this tutorial is **Bismark Bisulfite Mapper**. As with any software it is best to be famililar with the [User Manual](https://rawgit.com/FelixKrueger/Bismark/master/Docs/Bismark_User_Guide.html). Thus content below will provided with the assumption that the reader has read the manual and is meant to serve as guidance based on experience working with marine invertebrates.

The first step in the process in preparing the genome. Example code:

    # Directories and programs
    bismark_dir="/programs/Bismark-0.21.0"
    bowtie2_dir="/programs/bowtie2-2.3.4.1-linux-x86_64/"
    genome_folder="/where/the/fastafile/lives/"

    ${bismark_dir}/bismark_genome_preparation \
    --verbose \
    --parallel 28 \
    --path_to_aligner ${bowtie2_dir} \
    ${genome_folder}

You should expect to prepared genome with new directories including

    ./Bisulfite_Genome
    ./Bisulfite_Genome/GA_conversion
    ./Bisulfite_Genome/CT_conversion

For aligning the trimmed reads to the genome the following code structure is used

    find ${reads_dir}*_R1_001_val_1.fq.gz \
    | xargs basename -s _R1_001_val_1.fq.gz | xargs -I{} ${bismark_dir}/bismark \
    --path_to_bowtie ${bowtie2_dir} \
    -genome ${genome_folder} \
    -p 4 \
    -score_min L,0,-0.6 \
    --non_directional \
    -1 ${reads_dir}{}_R1_001_val_1.fq.gz \
    -2 ${reads_dir}{}_R2_001_val_2.fq.gz \
    -o Mcap_tg

This will create bam files (sequence alignment files)

Files are then deduplicated if they are whole genome bisulfite treated samples. This command will deduplicate the Bismark alignment BAM file and remove all reads that align to the the very same position and in the same orientation except for one. This step is recommended for whole-genome bisulfite samples, but **should not** be used for reduced representation libraries such as RRBS, amplicon, or target enrichment libraries. The [`bismark` manual](https://rawgit.com/FelixKrueger/Bismark/master/Docs/Bismark_User_Guide.html#ix-notes-about-different-library-types-and-commercial-kits) provides additional information.

    find *.bam | \
    xargs basename -s .bam | \
    xargs -I{} ${bismark_dir}/deduplicate_bismark \
    --bam \
    --paired \
    {}.bam

# Methylation Quantification

Methylation levels are then extracted using the `bismark_methylation_extractor`. For example:

    ${bismark_dir}/bismark_methylation_extractor \
    --bedGraph \
    --counts \
    --comprehensive \
    --merge_non_CpG \
    --multicore 28 \
    --buffer_size 75% \
    *deduplicated.bam

This will create a file with the suffix `deduplicated.bismark.cov.gz` (example shown below in uncompressed format). The resultant bedGraph file contains information from *unmerged* strands, meaning reads aligning to forward and reverse strands in the same position are not combined. Later in this workflow, a bedGraph with *merged* information will be created.

    NC_035784.1 141 141 37.5    3   5
    NC_035784.1 142 142 100 2   0
    NC_035784.1 155 155 70  7   3
    NC_035784.1 156 156 100 2   0
    NC_035784.1 291 291 0   0   2
    NC_035784.1 292 292 0   0   3
    NC_035784.1 313 313 0   0   1
    NC_035784.1 314 314 66.6666666666667    2   1
    NC_035784.1 470 470 66.6666666666667    4   2
    NC_035784.1 611 611 0   0   4

column organization of the file

    <chromosome> <start position> <end position> <methylation percentage> <count methylated> <count unmethylated>

A genome-wide cytosine report is generated (including merged bedGraph file), with this code as an example:

    find *deduplicated.bismark.cov.gz \
    | xargs basename -s _trimmed_bismark_bt2.deduplicated.bismark.cov.gz \
    | xargs -I{} ${bismark_dir}/coverage2cytosine \
    --genome_folder ${genome_folder} \
    -o {} \
    --merge_CpG \
    --zero_based \
    {}_trimmed_bismark_bt2.deduplicated.bismark.cov.gz

Generating a file `.CpG_report.merged_CpG_evidence.cov` (merged bedGraph file)

    NC_035785.1 217 219 100.000000  17  0
    NC_035785.1 523 525 87.500000   7   1
    NC_035785.1 556 558 50.000000   5   5
    NC_035785.1 727 729 100.000000  16  0
    NC_035785.1 1330    1332    0.000000    0   2
    NC_035785.1 1403    1405    0.000000    0   2
    NC_035785.1 1494    1496    66.666667   2   1
    NC_035785.1 1747    1749    100.000000  8   0
    NC_035785.1 2024    2026    100.000000  24  0
    NC_035785.1 2054    2056    93.333333   14  1

# File Conversions

From here it could be useful to do some simple file reorganizaion to obtain bedGraph or tab files for downstream analysis (eg bedtools, GLMs).

Creating bedGraphs post-merge at a specific coverage threshold (ex. 10x or 5x):

    for f in *merged_CpG_evidence.cov
    do
      STEM=$(basename "${f}" .CpG_report.merged_CpG_evidence.cov)
      cat "${f}" | awk -F $'\t' 'BEGIN {OFS = FS} {if ($5+$6 >= 10) {print $1, $2, $3, $4}}' \
      > "${STEM}"_10x.bedgraph
    done

    for f in *merged_CpG_evidence.cov
    do
      STEM=$(basename "${f}" .CpG_report.merged_CpG_evidence.cov)
      cat "${f}" | awk -F $'\t' 'BEGIN {OFS = FS} {if ($5+$6 >= 5) {print $1, $2, $3, $4}}' \
      > "${STEM}"_5x.bedgraph
    done

Creating tab files with raw count data for GLMs at 10x and 5x coverage:

    for f in *merged_CpG_evidence.cov
    do
      STEM=$(basename "${f}" .CpG_report.merged_CpG_evidence.cov)
      cat "${f}" | awk -F $'\t' 'BEGIN {OFS = FS} {if ($5+$6 >= 10) {print $1, $2, $3, $4, $5, $6}}' \
      > "${STEM}"_10x.tab
    done

    for f in *merged_CpG_evidence.cov
    do
      STEM=$(basename "${f}" .CpG_report.merged_CpG_evidence.cov)
      cat "${f}" | awk -F $'\t' 'BEGIN {OFS = FS} {if ($5+$6 >= 5) {print $1, $2, $3, $4, $5, $6}}' \
      > "${STEM}"_5x.tab
    done

It is sometimes useful to sort BAM files for downstream analysis (eg methylkit, IGV)

    # Sort files for methylkit and IGV
    find *.bam | \
    xargs basename -s .bam | \
    xargs -I{} ${samtools} \
    sort --threads 28 {}.bam \
    -o {}.sorted.bam

    # Index sorted files for IGV
    find *.sorted.bam | \
    xargs basename -s .sorted.bam | \
    xargs -I{} ${samtools} \
    index -@ 28 {}.sorted.bam

# Potential downstream applications

- Gene-level methylation analysis: people do this, but worth considering the utility of converting loci-level resolution data to gene-level resolution data
- methylKit
- IGV
- Analysis with other datasets (ex. pop gen, RNA-Seq)

# References