### 2023-12-15 DRAFT_lncRNA.Rmd
started Draft lncRNA discovery page

```{r,message = FALSE,echo = FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
library(knitcitations)
library(kableExtra)
opts_chunk$set(fig.width = 10,
               fig.height = 5,
               cache = FALSE)
cite_options(citation_format = "pandoc", max.names = 3, style = "html", hyperlink = "to.html")
```

---
title: 'Long non-coding RNA Discovery'
subtitle: '*Steven Roberts* and  *Zach Bengston*'
bibliography: DRAFT_lncRNA/DRAFT_lncRNA.bib
---

Initial publication year: TBD <br> [How to cite](https://marineomics.github.io/#How_to_Cite)

# Introduction

# References {#refs}