### 2023-12-15 DRAFT_lncRNA.Rmd
lncRNA - stringtie

```{r,message = FALSE,echo = FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
library(knitcitations)
library(kableExtra)
opts_chunk$set(fig.width = 10,
               eval = FALSE,         # Evaluate code chunks
               fig.height = 5,
               cache = FALSE,
               message = FALSE,
               warning = FALSE)
cite_options(citation_format = "pandoc", max.names = 3, style = "html", hyperlink = "to.html")
```

---
title: 'Long non-coding RNA Discovery'
subtitle: '*Steven Roberts* and  *Zach Bengston*'
bibliography: DRAFT_lncRNA/DRAFT_lncRNA.bib
---

Initial publication year: TBD <br> [How to cite](https://marineomics.github.io/#How_to_Cite)

# Introduction

Notes on running code. This workflow is written such that a user an fully reproduce the code.

# Downloading data

The data used for this workflow is from Acropora pulcra, and part of the E5 (https://e5coral.org) project. see https://github.com/urol-e5/deep-dive for more details.

```{r, engine='bash'}
# Make directories - these need to have a lot of space.
mkdir ../data/lncRNA
mkdir ../output/lncRNA
```

```{r, engine='bash'}
wget -r \
--no-directories --no-parent \
-P ../data/lncRNA \
-A "*fastq.gz" https://gannet.fish.washington.edu/Atumefaciens/20230519-E5_coral-fastqc-fastp-multiqc-RNAseq/A_pulchra/trimmed/

```

## Genome

For this we will be using the closly related genome on Acropora millipora.

```{r, engine='bash'}
cd ../data/lncRNA

curl -O https://gannet.fish.washington.edu/seashell/bu-github/deep-dive/D-Apul/data/Amil/ncbi_dataset/data/GCF_013753865.1/GCF_013753865.1_Amil_v2.1_genomic.fna
```

```{r, engine='bash'}
cd ../data/lncRNA

curl -O https://gannet.fish.washington.edu/seashell/bu-github/deep-dive/D-Apul/data/Amil/ncbi_dataset/data/GCF_013753865.1/genomic.gff
```

# Alignment

```{r, engine='bash'}
/home/shared/hisat2-2.2.1/hisat2-build \
../data/lncRNA/GCF_013753865.1_Amil_v2.1_genomic.fna \
../output/lncRNA/GCF_013753865.1_Amil_v2.1.index \
-p 48 \
../data/lncRNA/genomic.gff \
2> ../output/lncRNA/hisat2-build_stats.txt
```

```{r, engine='bash'}
find ../data/lncRNA/*R2_001.fastp-trim.20230519.fastq.gz \
| xargs basename -s -S1-TP2_R2_001.fastp-trim.20230519.fastq.gz | xargs -I{} \
/home/shared/hisat2-2.2.1/hisat2 \
-x ../output/lncRNA/GCF_013753865.1_Amil_v2.1.index \
-p 48 \
-1 ../data/lncRNA/{}-S1-TP2_R1_001.fastp-trim.20230519.fastq.gz \
-2 ../data/lncRNA/{}-S1-TP2_R2_001.fastp-trim.20230519.fastq.gz \
-S ../output/lncRNA/{}.sam \
2> ../output/lncRNA/hisat.out
```

# Convert BAMs to SAMs

```{r, engine='bash'}
for samfile in ../output/lncRNA/*.sam; do
  bamfile="${samfile%.sam}.bam"
  sorted_bamfile="${samfile%.sam}.sorted.bam"

  # Convert SAM to BAM
  /home/shared/samtools-1.12/samtools view -bS -@ 40 "$samfile" > "$bamfile"

  # Sort BAM
  /home/shared/samtools-1.12/samtools sort -@ 40 "$bamfile" -o "$sorted_bamfile"

  # Index sorted BAM
  /home/shared/samtools-1.12/samtools index -@ 40 "$sorted_bamfile"

  # remove sam files, no longer need
  rm *.sam

done

```

# Stringtie

```{r, engine='bash'}
find ../output/lncRNA/*sorted.bam \
| xargs basename -s .sorted.bam | xargs -I{} \
/home/shared/stringtie-2.2.1.Linux_x86_64/stringtie \
-p 16 \
-G ../lncRNA/genomic.gff \
-o ../output/lncRNA/{}.gtf \
../output/lncRNA/{}.sorted.bam
```

# References {#refs}